---
title: "Session 2: Import, Visualize, and Communicate"
author: "My Ha"
format: pdf
editor: visual
---

## Importing with Webscraping

Chapter 24 Web Scraping <https://r4ds.hadley.nz/webscraping.html#introduction>

To learn more about Quarto see <https://quarto.org.>

## Libraries needed

```{r}
library(rvest)
library(tidyverse)
```

```{r}
library(tibble)
library(purrr)
library(stringr)
library(readr)
library(readxl)
library(writexl)
library(ggplot2)
library(dplyr)
```

## Web scraping

Web scraping is possible because most pages that contain data that you want to scrape generally have a consistent structure.

\- elements

<https://developer.mozilla.org/en-US/docs/Web/HTML>

\- attributes

Two of the most important attributes are id and class, which are used in conjunction with CSS (Cascading Style Sheets) to control the visual appearance of the page.

##Target Site for webscraping

Top Trending Stocks: US stocks with the highest interest today - Yahoo Finance

<https://finance.yahoo.com/markets/stocks/trending/>

```{r}
url<- "https://finance.yahoo.com/markets/stocks/trending/"
html <- read_html(url)
```

Using Selectorgadget identify the desired table, select it, then copy the element .yf-1bczin to the statement below:

```{r}
section <- html |> html_elements(".yf-1bczin")
section
```

```{r}
html |> html_table()
```

```{r}
trending_table <- html |>
html_element("table") |>
html_table()
trending_table
```

## Check the datatypes of trending_table

```{r}
trending_table_datatypes <- tibble( column = names(trending_table),
type = map_chr(trending_table,
~ paste(class(.x), collapse = ", ")) )
print(trending_table_datatypes)
```

## Prepare the table for plotting

For this example we will move the table to Excel and clean the table there. We will learn more data operations later in the course.

write the trending_table to your working Directory as “trending_table_raw.xlsx”

```{r}
write_xlsx(trending_table,"trending_table_raw.xlsx")
```

## In Excel perform these changes to the data:

1.  Remove column 3

2.  Remove rows 2,6

3.  trim the price from column named “Price” and put in a column named “PRICE_REVISED”

4.  Change the datatype for Change to a number 4

5.  Change the datatype for Change % to a Percentage

6.  Delete the columns “Volume, Avg Vol (3M), Market Cap, P/E Ratio (TTM), 52 Wk Change %, 52 Wk Range”

    Rename the file to TRENDING_TABLE_REPAIRED.XLSX then load to your working directory

## Read the file into R

```{r}
Clean_Table <- read_xlsx("TRENDING_TABLE_REPAIRED.xlsx")
```

Change datatypes for Change, Change %, and PRICE_REVISED

```{r}
Clean_Table <- Clean_Table %>%
mutate(
Change = as.numeric(Change),
PRICE_REVISED = as.numeric(PRICE_REVISED),
`Change %` = str_remove(`Change %`, "%"), # remove %
`Change %` = as.numeric(`Change %`) # convert to decimal
)
```

## Here is where we begin our journey with ggplot()

ggplot(): <https://rstudio.github.io/cheatsheets/html/data-visualization.html>

Layers: ggplot(data = ) + (mapping = aes(), stat = , position = ) + + + +

Cheat Sheet : <https://rstudio.github.io/cheatsheets/data-visualization.pdf>

## Now create a column chart that demonstrates the stock prices for each company

```{r}
Original <- ggplot(Clean_Table, aes(x = reorder(Name, PRICE_REVISED),
y = PRICE_REVISED)) +
geom_col(fill = "steelblue") +
coord_flip() +
labs( title = "Price (PRICE_REVISED) by Company",
x = "Company", y = "Price (Revised)" ) + theme_minimal(base_size = 12)
print(Original)
```

## Clean the plot with the changes below and print to a new name

```{r}
Cleaner_view <- ggplot(
  Clean_Table,
  aes(x = reorder(Name, PRICE_REVISED), y = PRICE_REVISED)
) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Trending Price by Company",
    x = "Company",
    y = "Trending Price"
  ) +
  theme_minimal(base_size = 14)

print(Cleaner_view)
```

## Create a scatter plot to show Change % and Price by Name

```{r}
Scatter <- ggplot(
  Clean_Table,
  aes(x = `Change %`, y = PRICE_REVISED, label = Name)
) +
  geom_point(color = "steelblue", size = 6) +
  geom_text(nudge_y = 1, size = 3.2) +
  labs(
    title = "Scatter Plot of Change % vs Price (Revised)",
    x = "Change (%)",
    y = "Price"
  ) +
  theme_minimal(base_size = 8)

print(Scatter)
```

When you click the Render button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

The echo: false option disables the printing of code (only output is displayed)
