---
title: "Section 6: Non-Linear Regression Modeling"
author: "My Ha"
format: html
editor: visual
---

## Linear Regression Modeling in R

Section: 02

Name: My Ha

###References

To learn more about Quarto see <https://quarto.org>

#### Underlying Non-linear Equations for this analysis:

Sphere Radius = (X^2+y^2+z^2)^.5

Square Area = X \* Y

Ring Area = pi()*R\^2-pi()*r\^2

Cube Volume = X \* Y \* Z

Cylindrical Volume = pi()*R\^2*h

Sphere Volume = 4*pi()*R\^2/3

### Install Packages- comment out when complete

```{r}
#install.packages(c("rlang", "vctrs", "pillar", "tibble", "dplyr","plotly"))
```

### Install Libraries

```{r}
library(tidyverse)
```

```{r}
library(readxl)
library(reshape2)
library(plotly)
```

### Set Working Directory and load file into that location

```{r}
getwd()
```

## Read the data from an excel Spreadsheet. We will test this data to see how well it fits to the prescribed formulas.

```{r}
df <- read_excel("data for modeling.xlsx")
```

### Exploratory Data Analysis of All Variables - Ensuring the data is capable of analysis.

```{r}
head(df)
str(df)
summary(df)
```

### Distributions - Observation of the distributions is key to proper usage of a variables

[Question #1: Which Distribution appears to be skewed right more than the others?]{.underline}

Answer: Cylinder_Volume, Sphere_Volume

The distributions for **Cylinder_Volume** and **Sphere_Volume** appear to be the most right-skewed among all variables. This is evident from their long right tails, where a small number of observations have significantly larger values compared to the rest

```{r}
df %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  ggplot(aes(x = Value)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  facet_wrap(~ Variable, scales = "free") +
  theme_minimal()
```

### Pairwise Relationships

[Question #2: Which Pairwise comparison(s) seem to more strongly correlated?]{.underline}

Answer: Sphere_Volume ↔ Ring, Z ↔ R, Z ↔ r

The strongest pairwise correlations are observed between **Sphere_Volume and Ring**, as well as **Z and R**, and **Z and r**, as these pairs show clear linear patterns with tightly clustered points. These relationships are strong because the variables are mathematically related through geometric formulas.

```{r}
pairs(df)
```

### Correlation Matrix

```{r}
print(cor(df))
```

### Outlier Detection

[Question #3: Why do “Sphere Volume” and “Cylinder Volume” appear to have more outliers than the others?]{.underline}

Answer: **Sphere_Volume** and **Cylinder_Volume** appear to have more outliers because they are calculated using squared and cubic relationships involving the radius. As a result, small increases in radius lead to disproportionately large increases in volume. This amplifies extreme values and creates a wider spread in the data, making high-end observations appear as outliers compared to other variables.

```{r}
df %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  ggplot(aes(x = Variable, y = Value)) +
  geom_boxplot(fill = "lightgreen") +
  facet_wrap(~ Variable, scales = "free") +
  theme_minimal() +
  theme(axis.text.x = element_blank())
```

### Non-Linear regressions

#### Quadratic Regression of X on X\^2

```{r}
model_x2 <- lm(X^2 ~ X, data = df)
summary(model_x2)
```

```{r}
ggplot(df, aes(x = X, y = X^2)) +
  geom_point(color = "steelblue", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Quadratic Regression: X^2 on X", x = "X", y = "X^2") +
  theme_minimal()
```

#### Quadratic regression of X, Y, and X\*Y on Square_Area

#####Square Area by X only

```{r}
model_sq_x <- lm(Square_Area ~ X, data = df)
summary(model_sq_x)
```

```{r}
ggplot(df, aes(x = X, y = Square_Area)) +
  geom_point(color = "steelblue", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Square Area by X", x = "X", y = "Square_Area") +
  theme_minimal()
```

#####Square Area by Y only

```{r}
model_sq_y <- lm(Square_Area ~ Y, data = df)
summary(model_sq_y)
```

```{r}
ggplot(df, aes(x = Y, y = Square_Area)) +
  geom_point(color = "steelblue", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Square Area by Y", x = "Y", y = "Square_Area") +
  theme_minimal()
```

[Question #4: Why do X and Y so poorly explain Square_Area?]{.underline}

Answer: Square_Area is defined as X × Y, meaning it depends on both variables together in a multiplicative (nonlinear) way. When you regress Square_Area on X alone, you ignore Y, which is a key part of what determines the area. Because X and Y vary independently, neither variable by itself captures this joint effect. As a result, each model explains only a limited portion of the variation, keeping R² below 40%.

#####Since X nor Y explain the Square area well (\<40% R\^2) we need to look elsewhere like by X\*Y

```{r}
df <- df %>% mutate(XY = X * Y)
model_sq_xy <- lm(Square_Area ~ XY, data = df)
summary(model_sq_xy)
```

```{r}
ggplot(df, aes(x = XY, y = Square_Area)) +
  geom_point(color = "steelblue", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Square Area by X*Y", x = "X*Y", y = "Square_Area") +
  theme_minimal()
```

##### X\*Y does explain the Square Area better with a low p-value (as did X and Y) and twice the correlation

[Question #5 : Why does X\*Y better explain Square_Area?]{.underline}

Answer:

##### Cube Volume by X*Y*Z

```{r}
df <- df %>% mutate(XYZ = X * Y * Z)
model_cube <- lm(Y ~ XYZ, data = df)
summary(model_cube)
```

```{r}
ggplot(df, aes(x = XYZ, y = Cube_Volume)) +
  geom_point(color = "steelblue", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Cube Volume by X*Y*Z", x = "X*Y*Z", y = "Cube_Volume") +
  theme_minimal()
```

##### While XYZ are important to the Cube Volume (low p=value) there is still variation which is not understood

##### R by X, Y, and Z

```{r}
df <- df %>% mutate(Rcalc = sqrt(X^2 + Y^2 + Z^2))
model_R <- lm(R ~ Rcalc, data = df)
summary(model_R)
```

```{r}
ggplot(df, aes(x = Rcalc, y = R)) +
  geom_point(color = "steelblue", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "R by Calculated Sphere Radius", x = "Rcalc = sqrt(X^2+Y^2+Z^2)", y = "R") +
  theme_minimal()
```

##### While the formula for the radius of a sphere is important for describing R (low p=value) there is still some variation which is not understood (70% R\^2)

##### Cylindrical Volume by R and h

```{r}
df <- df %>% mutate(CylVolume = pi * R^2 * h)
model_cyl <- lm(Cylinder_Volume ~ CylVolume, data = df)
summary(model_cyl)
```

```{r}
ggplot(df, aes(x = CylVolume, y = Cylinder_Volume)) +
  geom_point(color = "steelblue", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Cylinder Volume by pi*R^2*h", x = "CylVolume (calculated)", y = "Cylinder_Volume") +
  theme_minimal()
```

##### The quadratic equation used to represent Cyl_Volume poorly explains the Cylinder_Volume due to the high p=value and low correlation. This tells use there is still variation which is not understood.

##### Spherical Volume by R

```{r}
df <- df %>% mutate(SpherVolume = (4 * pi * R^3) / 3)
model_sph <- lm(Sphere_Volume ~ SpherVolume, data = df)
summary(model_sph)
```

```{r}
ggplot(df, aes(x = SpherVolume, y = Sphere_Volume)) +
  geom_point(color = "steelblue", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Sphere Volume by (4/3)*pi*R^3", x = "SpherVolume (calculated)", y = "Sphere_Volume") +
  theme_minimal()
```

##### The cubic equation used to represent Spher_Volume poorly explains the Sphere_Volume due to the high p=value and low correlation. This tells use there is still variation which is not understood.

##### Ring by R

```{r}
df <- df %>% mutate(Ring_area = pi * R^2 - pi * r^2)
model_ring <- lm(Ring ~ Ring_area, data = df)
summary(model_ring)
```

```{r}
ggplot(df, aes(x = Ring_area, y = Ring)) +
  geom_point(color = "steelblue", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Ring by pi*R^2 - pi*r^2", x = "Ring_area (calculated)", y = "Ring") +
  theme_minimal()
```

##### The compound quadratic equation used to represent Ring_area poorly explains Ring due to the high p=value and low correlation. This tells use there is still variation which is not understood.
