---
title: "Section 5 Data Tidying Assignment"
author: "My Ha"
format: pdf
editor: visual
---

## Session 5: Tidying Data Assignment

Section: 02

Name: My Ha

## References

To learn more about Quarto see <https://quarto.org>

Course Text Link:

[Chapter 5: R for Data Science](https://r4ds.hadley.nz/data-tidy.html)

### Install Packages- comment out when complete

```{r}
#install.packages(c("rlang", "vctrs", "pillar", "tibble", "dplyr"))
```

## Install Libraries

```{r}
library(tidyverse)
```

```{r}
library(readxl)
library(ggplot2)
library(reshape2)
```

## Set Working Directory and load file into that location

```{r}
getwd()
```

## Read in data for assignment

```{r}
Section5data <-read_xlsx("Section5_Data_Tidying_Assignment.xlsx")
```

## RECALL: What makes a dataset a “tidy” one

### There are three interrelated rules that make a dataset tidy:

1\. Each variable is a column; each column is a variable.

2\. Each observation is a row; each row is an observation.

3\. Each value is a cell; each cell is a single value

### review the data in Section5data and notice the opportunities to tidy the data

```{r}
Section5_Pivot_longer <- Section5data |>
pivot_longer(
cols = c("Satisfied","Neither","Dissatisfied"), #columns to be pivotted
names_to = "Survey_Result", # new column name
values_to = "Total_Quantity" # values in new column,
)
```

## Visualize the results

```{r}
ggplot(Section5_Pivot_longer, aes(x = Survey_Result, y = Total_Quantity)) +
geom_boxplot(fill = "#E67E22", alpha = 0.7) +
labs(
title = "Total Quantity by Survey Result",
x = "Survey Result",
y = "Total Quantity"
) +
theme_minimal(base_size = 14)
```

```{r}
Satisfied <- Section5_Pivot_longer %>% filter(Survey_Result == "Satisfied")
Neither <- Section5_Pivot_longer %>% filter(Survey_Result == "Neither")
Dissatisfied <- Section5_Pivot_longer %>% filter(Survey_Result == "Dissatisfied")
```

```{r}
ggplot(Satisfied, aes(x = OrderDate, y = Total_Quantity)) +
geom_line(color = "#2C3E50", linewidth = 1) +
geom_point(color = "#2980B9", size = 2) +
labs(
title = "Total Quantity Over Time",
x = "Order Date",
y = "Total Quantity"
) +
theme_minimal(base_size = 14)
```

```{r}
ggplot(Satisfied, aes(x = Total_Quantity)) +
geom_histogram(aes(y = after_stat(density)),
bins = 10, fill = "#3498DB", alpha = 0.6) +
geom_density(color = "#2C3E50", linewidth = 1) +
labs(
title = "Distribution of Satisfied",
x = "Total Quantity",
y = "Density"
) +
theme_minimal(base_size = 14)
```

```{r}
ggplot(Neither, aes(x = Total_Quantity)) +
geom_histogram(aes(y = after_stat(density)),
bins = 10, fill = "#3498DB", alpha = 0.6) +
geom_density(color = "#2C3E50", linewidth = 1) +
labs(
title = "Distribution of Neither",
x = "Total Quantity",
y = "Density"
) +
theme_minimal(base_size = 14)
```

```{r}
ggplot(Dissatisfied, aes(x = Total_Quantity)) +
geom_histogram(aes(y = after_stat(density)),
bins = 10, fill = "#3498DB", alpha = 0.6) +
geom_density(color = "#2C3E50", linewidth = 1) +
labs(
title = "Distribution of Dissatisfied",
x = "Total Quantity",
y = "Density"
) +
theme_minimal(base_size = 14)
```

```{r}
ggplot(Section5_Pivot_longer, aes(x = Total_Quantity, fill = Survey_Result)) +
geom_density(alpha = 0.5, color = "black") +
facet_wrap(~ Survey_Result, ncol = 2) +
labs(
title = "Density of Total Quantity by Survey Result",
x = "Total Quantity",
y = "Density"
) +
theme_minimal(base_size = 14) +
theme(legend.position = "none")
```
